---
title: "Stat 3302 Project"
author: "Yoojin Jeong"
date: "4/6/2022"
output: pdf_document
---

Loading the dataset in R
```{r}
library(readr)
diamonds <- read_csv("diamonds.csv")
View(diamonds)
```


Linear regression for price and carat size
```{r}
diamonds.lm = lm(price ~ carat, data = diamonds)
plot(diamonds$carat, diamonds$price, xlab = "Carat Size", ylab = "Price ($)")
abline(reg = diamonds.lm)

#Summary statistics
summary(diamonds.lm)
```


From the above linear regression and plot, it appears that fitting price and carat size to a linear regression model will not work.



Taking Log Predictor and Response to fit model into linear regression- INCLUDE IN REPORT
```{r}
#Log Predictor
plot(log(diamonds$carat), diamonds$price, xlab = "Carat Size", ylab = "Price ($)")
abline(reg = diamonds.lm)
#Log Response
plot(diamonds$carat, log(diamonds$price), xlab = "Carat Size", ylab = "Price ($)")
abline(reg = diamonds.lm)

#Log Predictor AND Log Response
diamonds.lm.log = lm(log(price) ~ log(carat), data = diamonds) 

plot(log(diamonds$carat), log(diamonds$price), xlab = "Carat Size", ylab = "Price ($)", main = "Price vs Carat Size")
abline(reg = diamonds.lm.log, col = "red")
summary(diamonds.lm.log)
```


Analysis: 


Cut as Factor Variable
```{r}
model.fac <- glm(price ~ carat + as.factor(cut), data = diamonds)
#Cut "Fair' is baseline
summary(model.fac)


diamonds.df <- data.frame(diamonds, stringsAsFactors = TRUE)
model.fac2 <- glm(price ~ carat + as.factor(cut) + diamonds$color, data = diamonds)
summary(model.fac2)
```



Multivariate multiple regression.
```{r}
multi.reg <- lm(price ~ carat + depth, data = diamonds)
summary(multi.reg)

multi.reg2 <- lm(price ~ carat + depth + table, data = diamonds)
summary(multi.reg2)

anova(multi.reg, multi.reg2)
```

Proportional Odds Model Regression
```{r}
library(MASS)
hist(diamonds$price)
plot(diamonds$carat,diamonds$price)

diamonds$pricerange="0-4999"
for (i in length(diamonds$price)) {
  if(diamonds$price[i]>4999) diamonds$pricerange[i]="5000-9999"
}
for (i in length(diamonds$price)) {
  if(diamonds$price[i]>9999) diamonds$pricerange[i]="10000-14999"
}
for (i in length(diamonds$price)) {
  if(diamonds$price[i]>14999) diamonds$pricerange[i]="15000+"
}
diamonds$pricerange=cut(diamonds$price, breaks = c(0,5000,10000,15000,20000), labels = c("0-4999", "5000-9999", "10000-14999", "15000+"))
diamonds$pricerangef=factor(diamonds$pricerange, levels = c("0-4999", "5000-9999", "10000-14999", "15000+"))
summary(polr(diamonds$pricerangef~diamonds$carat))
```

